Large and Infinitary Quotient Inductive-Inductive Types

### 2020-04-12 draft/paper.tex
fmt change & induction response

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[sigplan]{acmart}\settopmatter{}

@@

\author{Andr{\'a}s Kov{\'a}cs}
\orcid{0000-0002-6375-9781}
\affiliation{
  \department{Department of Programming Languages and Compilers}
  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
  \city{Budapest}
  \country{Hungary}                   %% \country is recommended
}
\email{kovacsandras@inf.elte.hu}         %% \email is recommended

@@

A prime motivation of the current work is to further develop QIITs as a
framework for the metatheory of type theories, to cover more theories and
support more applications. To this end, we extend the syntax and semantics of
QIITs as previously described in the literature \cite{kaposi2019constructing,
  qiits, dijkstra2017quotient}.

@@

\begin{definition}[Finite limit CwFs]\label{def:flCwF}
For each level $i$ we define $\flCwF_i : \Set_{i+1}$ as an iterated $\Sigma$-type
with the following components:
\begin{enumerate}
  \item A CwF with underlying sets all in $\Set_i$. We reuse the component
    notations $\Con$, $\Sub$, $\Ty$, etc.\ from Definition \ref{def:tos}.
  \item $\Sigma$-types $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
    \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
  \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with $\refl$, equality reflection and UIP.
  \item Constant families. This includes a type former $\Kfam : \Con \ra
    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, together with $\lamK
    : \Sub\,\Gamma\,\Delta \ra \Tm\,\Gamma\,(\Kfam\,\Delta)$ and its inverse
    $\appK$. The idea is that $\Kfam\,\Delta$ is a representation of $\Delta$ as
    a type in any context. Clairambault and Dybjer called constant families
    ``democracy'' in \cite{clairambault2014biequivalence}.
\end{enumerate}
\end{definition}

\begin{definition} We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
$\Kfam$ as \emph{fl-structure}.
\end{definition}

@@

In the categorical semantics for a concrete signature, this informally means
that an algebra supports induction if for any dependent algebra over it (which
is a bundle of induction motives and methods), there is a dependent morphism
into it (which is a bundle of eliminator functions and their $\beta$-rules).

@@

In the following, we will need three model fragments of $\bM$, which can be used
to compute notions of algebras, displayed algebras and sections respectively for
each syntactic signature. This is a rephrasing of the $\blank^{\A}$,
$\blank^{\D}$ and $\blank^{\S}$ interpretations in
\cite{kaposi2019constructing}.

>>>>

\documentclass[sigplan,screen]{acmart}

@@

\author{Andr{\'a}s Kov{\'a}cs}
\orcid{0000-0002-6375-9781}
\affiliation{
  %% \department{Department of Programming Languages and Compilers}
  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
  \city{Budapest}
  \country{Hungary}
}
\email{kovacsandras@inf.elte.hu}

\author{Ambrus Kaposi}
\orcid{0000-0001-9897-8936}
\affiliation{
  %% \department{Department of Programming Languages and Compilers}
  \institution{E{\"o}tv{\"o}s Lor{\'a}nd University}
  \city{Budapest}
  \country{Hungary}
}
\email{akaposi@inf.elte.hu}

@@

A prime motivation of the current work is to further develop QIITs as a
framework for the metatheory of type theories, to cover more theories and
support more applications. To this end, we extend the syntax and semantics of
QIITs as previously described in the literature \cite{kaposi2019constructing,
  qiits, dijkstra2017quotient}, with the following features.

@@

\begin{definition}[Finite limit CwFs]\label{def:flCwF}
For each level $i$ we define $\flCwF_i : \Set_{i+1}$ as an iterated $\Sigma$-type
with the following components:
\begin{enumerate}
  \item A CwF with underlying sets all in $\Set_i$. We reuse the component
    notations $\Con$, $\Sub$, $\Ty$, etc.\ from Definition \ref{def:tos}.
  \item $\Sigma$-types $\Sg : (A : \Ty\,\Gamma)\ra \Ty\,(\Gamma \ext A)
    \ra \Ty\,\Gamma$, with term formers $\proj1$, $\proj2$ and $\blank,\blank$.
  \item Identity type $\Id : (A : \Ty\,\Gamma)\ra \Tm\,\Gamma\,A\ra
    \Tm\,\Gamma\,A\ra \Ty\,\Gamma$, with $\refl$, equality reflection and UIP.
  \item Constant families. This includes a type former $\Kfam : \Con \ra
    \Ty\,\Gamma$, where $\Gamma$ is implicitly quantified, together with $\lamK
    : \Sub\,\Gamma\,\Delta \ra \Tm\,\Gamma\,(\Kfam\,\Delta)$ and its inverse
    $\appK$. The idea is that $\Kfam\,\Delta$ is a representation of $\Delta$ as
    a type in any context. Clairambault and Dybjer called constant families
    ``democracy'' in \cite{clairambault2014biequivalence}.
\end{enumerate}
We abbreviate the additional structure on CwFs consisting of $\Sigma$, $\Id$ and
$\Kfam$ as \emph{fl-structure}.
\end{definition}

@@

For an example, if we interpret $\mathsf{\NatSig}$ in the $\bM$ model, we get an
flCwF of natural number algebras, where $\Con$ is the type of algebras and
$\Sub\,\Gamma\,\Delta$ is the type of homomorphisms between $\Gamma$ and
$\Delta$ algebras. $\Ty$ is the type of displayed algebras, and $\Tm$ is the
type of their sections:
\begin{alignat*}{3}
  & \Ty\,(N,\,z,\,s) \equiv(N^D : N \to \Set)\\ & \hspace{2em}\times (N^D\,z)
  \times ((n : N)\to N^D\,n \to N^D(s\,n))\\ &
  \Tm\,(N,\,z,\,s)\,(N^D,\,z^D,\,s^D) \equiv (N^S : (n : N)\to
  N^D\,n)\\ & \hspace{2em}\times (N^S\,z = z^D)\times ((n : N) \to N^S\,(s\,n) =
  s^D\,n\,(N^S\,n))
\end{alignat*}
Thus, for natural number algebras, $\Inductive$ is exactly the predicate which
holds when an algebra supports induction.

%% $\Inductive$ can be viewed as a
%% dependently typed analogue of initiality: an algebra $\Gamma$ is inductive if
%% for every dependent algebra $A : \Ty\,\Gamma$ over it there is a dependent
%% morphism from $\Gamma$ to $A$, given as an element of $\Tm\,\Gamma\,A$.

%% For some $(N,\,z,\,s)$ algebra,
%% $\Ty\,(N,\,z,\,s)$ is defined as $(N^P : N \to \Set)\times N^P\,z \times ((n :
%% N)\to N^P\,n \to N^P(s\,n))$, the type of displayed algebras, and
%% $\Tm\,(N,\,z,\,s)\,(N^P,\,z^P,\,s^P)$ is $(N^S : (n : N)\to N^P\,n) \times (N^S\,z = z^P)
%% \times ((n : N) \to N^S\,(s\,n) = s^P\,(N^S\,n))$.

%% Here, $\Con_{\bGamma}$ is the type of algebras for some signature, e.g.\ the
%% type of natural number algebras in the case of
%% $\mathsf{NatSig}$. $\Ty_{\bGamma}\,\Gamma$ is a displayed algebra over a
%% concrete algebra

%% In the categorical semantics for a concrete signature, this informally means
%% that an algebra supports induction if for any dependent algebra over it (which
%% is a bundle of induction motives and methods), there is a dependent morphism
%% into it (which is a bundle of eliminator functions and their $\beta$-rules).

@@

In the following, we will need three model fragments of $\bM$, which can be used
to compute notions of algebras, displayed algebras and sections respectively for
each syntactic signature. This is a rephrasing of the $\blank^{\A}$,
$\blank^{\D}$ and $\blank^{\S}$ interpretations in
\cite{kaposi2019constructing}, where they are discussed at more length.
